---
title: "Day 1: Intro to **R** and ```tidyverse```"
output:
  html_notebook:
    highlight: haddock
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---



```{r setup, include = FALSE} 
knitr::opts_chunk$set(comment = '#',
                      tidy = TRUE)
```



## Setup

If you want a fresh start:

* clear plots (if no plots are present, comment it out or it will throw an error)

```{r setup_clear_plots, eval = FALSE}
dev.off()
```

* clear console

```{r setup_clear_console, eval = FALSE}
cat("\014")
```

* clear environment

```{r setup_clear_environment, eval = FALSE}
rm(list = ls())
```

How to install packages:

```{r setup_install_packages, results  = 'hide'}
install.packages("tidyverse")
```

It's good to know where your packages are. Type ```.libPaths()```.

```{r setup_install_packages_dir, echo = FALSE}
.libPaths()
```

How to load packages:

```{r setup_load_packages, results  = 'hide'}
library(tidyverse)
```

We are now inside an **R** project, which means that the working directory is automatically set to the directory containing the current *.Rmd* file. We can verify it by typing ```getwd()```.

```{r setup_where_are_we, echo = FALSE}
getwd()
```

This is my path, yours will obviously differ. If, for whatever reason, you are in another directory, you can set the working directory with ```setwd()```.

```{r setup_where_should_you_be, eval = FALSE}
setwd("NAME/OF/YOUR/DIRECTORY")
```

Verify that you are in the right directory with ```getwd()```. If you still have problems, please ask for help.

## The data

### Data description

The data come from 

There is evidence from advertising research that attitudes towards stimuli can be changed using
positive imagery (e.g., Stuart, Shimp, & Engle, 1987). As part of an initiative to stop binge drinking in
teenagers, the government funded some scientists to look at whether negative imagery could be used
to make teenagers’ attitudes towards alcohol more negative. The scientists designed a study to address
this issue by comparing the effects of negative imagery against positive and neutral imagery for
different types of drinks. Table 13.4 illustrates the experimental design and contains the data for this
example (each row represents a single participant).
Participants viewed a total of nine mock adverts over three sessions. In one session, they saw three
adverts: (1) a brand of beer (Brain Death) presented with a negative image (a dead body with the
slogan ‘drinking Brain Death makes your liver explode’); (2) a brand of wine (Dangleberry) presented
in the context of a positive image (a sexy naked man or woman – depending on the participant’s
preference – and the slogan ‘drinking Dangleberry wine makes you irresistible’); and (3) a brand of
water (Puritan) presented alongside a neutral image (a person watching television accompanied by the
slogan ‘drinking Puritan water makes you behave completely normally’). In a second session (a week
later), the participants saw the same three brands, but this time Brain Death was accompanied by the
positive imagery, Dangleberry by the neutral image and Puritan by the negative. In a third session, the
participants saw Brain Death accompanied by the neutral image, Dangleberry by the negative image
and Puritan by the positive. After each advert participants were asked to rate the drinks on a scale
ranging from – 100 (dislike very much) through 0 (neutral) to 100 (like very much). The order of
adverts was randomized, as was the order in which people participated in the three sessions. This
design is quite complex. There are two independent variables: the type of drink (beer, wine or water)
and the type of imagery used (positive, negative or neutral). These two variables completely cross
over, producing nine experimental conditions

### Load the data

The data are in *.dat* format, so we are going to use ```read.table```.

```{r load_data}
attitude <- 
    read.table(
        "MixedAttitude.dat", # file name
        header = TRUE)       # data has headers
```

If you don't know what a function does, access the help (e.g., ```?read.table```).   
   
Let's have a look at the data. If you have **RStudio**, you can click on the name of the dataset in the *Environment* window (by default on the upper right corner).   
   
You can also type the name of the dataset in the console:

```{r see_data}
attitude
```

### Clean & save the data













